{\small
\noindent
{\em Software:} Copyright \copyright\ 1978-2019 Gary J. Ferland and others. All rights reserved.

\vspace{1em}

\noindent
The software described in this documentation (\Cloudy) is subject
to a FreeBSD-style software license
contained in the file \cdFilename{license.txt} in the
root directory of the distributed files.
The list of co-authors is given in the file
\cdFilename{others.txt} in the same directory.
Use of this program is not restricted provided each use is
acknowledged upon publication.
The bibliographic reference to this version of \Cloudy\ is
``version xx.xx
of the code last described \citet{CloudyReview}.''
The version number, shown here as ``xx.xx'', should be
given.
This version number, along with a complete citation,
can be found by executing the code with the 
\cdCommand{print citation} command included in the input script.

\vspace{1em}

\noindent
\Cloudy\ is an evolving code.
You should confirm that you have the most recent
version of the code by checking the web site
\href{http://www.nublado.org}{www.nublado.org}.
The code has a
\href{http://tech.groups.yahoo.com/group/cloudy_simulations}{discussion board} with emailing list.
This will have announcements of any updates to the code.\par

\vspace{1em}

\noindent
Portions of the documentation have been published, and are copyrighted by, the
American Astronomical Society, the Astronomical Society of the Pacific, and
the Royal Astronomical Society. The remainder of the documentation is subject
to the following FreeBSD format documentation license:

\vspace{1em}

\noindent
{\em Documentation:} Copyright \copyright\ 1978--2019 Gary J. Ferland and others. All rights reserved.

\vspace{1em}

\noindent
Redistribution and use of the documentation (all parts of \Hazy\ and the
Quick Start Guide) in source (\LaTeX) and `compiled' forms (PDF, PostScript,
HTML and so forth) with or without modification, are permitted provided that
the following conditions are met:

\begin{enumerate}
\item
Redistributions of source code (\LaTeX) must retain the above copyright notice,
this list of conditions and the following disclaimer as the first lines of
the file \cdFilename{license\_doc.txt} in the root directory unmodified.
\item
Redistributions in compiled form (converted to PDF, PostScript, XML, HTML and
other formats) must reproduce the above copyright notice, this list of
conditions and the following disclaimer in the documentation and/or other
materials provided with the distribution.
\end{enumerate}

\noindent
{\em This documentation is provided by the \Cloudy\ project ``as is'' and any express
or implied warranties, including, but not limited to, the implied warranties
of merchantability and fitness for a particular purpose are disclaimed. In no
event shall the  \Cloudy\  project be liable for any direct, indirect, incidental,
special, exemplary, or consequential damages (including, but not limited to,
procurement of substitute goods or services; loss of use, data, or profits; or
business interruption) however caused and on any theory of liability, whether
in contract, strict liability, or tort (including negligence or otherwise)
arising in any way out of the use of this documentation, even if advised of
the possibility of such damage.}\par
}
